---
title: "SNP Visualization"
output: html_notebook
---
SNP Visualization: This R script visualizes SNP data from the gnomAD website. The four graphs shows 
1.The relative location of the parental gene of the ORF of interest
2.Relative location of SNPs and ORFs, including the ORF of interest, on the parental gene
3.The specific DNA sequences of the exons of the ORF of interest corresponding to the SNPs
4.Protein consequence and strand direction if further SNP data is provided

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("ensembldb", force = T)
install.packages("locuszoomr")
BiocManager::install("AnnotationHub")
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(purrr)
library(AnnotationHub)
library(locuszoomr)
library(biomaRt)
library(patchwork)
install.packages("ggbreak")
library(ggbreak) 
library(ensembldb)
library("Biostrings")

#Load ensembl data
ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

ah <- AnnotationHub()
query(ah, c("EnsDb", "Homo sapiens"))
ensDb_v106 <- ah[["AH100643"]]
```

The zoom_out_track function graphs the parental gene of the ORF of interest and its surrounding genes
Input: 
- Name of the specific ORF of a lncRNA from the Ribo-seq_ORFs.csv file 

Output: 
- Gene tracks of the parental gene and other genes within 100,000 base pairs of the parental gene with genomic position as the X-axis
- The parental gene is colored red

```{r}

zoom_out_tracks <- function(orf_name){
  gene_name <- orf %>%
    filter(name2 == orf_name) %>%
    pull(geneName2)
  test_position <- data.frame(chromosome = lollipop$chrom, position = lollipop$pos, p = lollipop$AF, rsId = lollipop$orf_id)
  
  loci <- locuszoomr::locus(data = test_position, gene = gene_name, flank = 1e5, ens_db = ensDb_v106)
  locuszoomr::genetracks(loci, highlight = gene_name)
}

```

The track_with_snp function creates a lollipop plot of the SNPs in all the ORFs in the parental gene of the ORF of interest. It is placed on top of the gene track of the parental gene
Input: 
- Name of the specific ORF of a lncRNA from the Ribo-seq_ORFs.csv file 

Output: 
- Lollipop plot of the SNPs downloaded from gnomAD with the allele frequency as the Y-axis and genomic position as the X-axis. The SNPs are distinguished by different colors based on their classification as SNV, insertion, deletion, duplication, delin, or other
- The exons of the ORF of interest is indicated by tall black boxes while the exons of other ORFs in the parental gene is indicated by small black boxes below the X-axis
- The gene track of the parental gene is directly below the lollipop plot

```{r}

track_with_snp <- function(orf_name){ 
  gene_name <- orf %>%
    filter(name2 == orf_name) %>%
    pull(geneName2)
  gene_symbol <- gene_name

  # Retrieve genomic coordinates for the gene
  gene_coordinates <- getBM(
      attributes = c("ensembl_gene_id", "external_gene_name", "chromosome_name", "start_position", "end_position"),
      filters = "hgnc_symbol",
      values = gene_symbol,
      mart = ensembl
  )
  
  gene_start <- as.numeric(gene_coordinates$start_position)
  gene_end <- as.numeric(gene_coordinates$end_position)
  
  linc_orf <- orf[orf$geneName2 == gene_name, ]

  rect <- data.frame(chromstart = linc_orf$chromStart, count = linc_orf$blockCount, size = linc_orf$blockSizes, start = linc_orf$blockStart)
  
  # Convert size and start columns to numeric vectors
  orf_df <- rect %>%
    mutate(size = map(strsplit(as.character(size), ","), function(x) as.numeric(x)),
           start = map(strsplit(as.character(start), ","), function(x) as.numeric(x)))
  
  # Function to create multiple columns based on count
  create_orf_columns <- function(chromstart, count, start, size) {
    # Initialize empty list to store results
    results <- list()
    
    # Loop over each count
    for (i in seq_len(count)) {
      # Calculate orfstart and orfend for each count
      orfstart <- chromstart + start[i]
      orfend <- chromstart + start[i] + size[i]
      
      # Store results in a list
      results[[paste0("orfstart", i)]] <- orfstart
      results[[paste0("orfend", i)]] <- orfend
    }
    
    # Return the results as a data frame
    return(as.data.frame(results))
  }
  

  
  # Apply the function row-wise using pmap from purrr
  orf_df <- bind_cols(orf_df, pmap_df(orf_df, create_orf_columns))
  orf_df$name <- linc_orf$name
  orf_df <- orf_df %>%
    mutate(ymax = ifelse(!is.null(orf_name) & name == orf_name, Inf, 0))
  orf_df$chromEnd <- linc_orf$chromEnd
  
  iterations <- 8 - orf_df$count
    
  for (i in (orf_df$count + 1):iterations) {
    orf_df[[paste0("orfstart", i)]] <- 0
    orf_df[[paste0("orfend", i)]] <- 0
  }
  
  new <- orf %>%
    filter(geneName2 == gene_name) %>%
    pull(geneName)
  
  filtered_lollipop <- lollipop %>%
    filter(gene_id == new)
  
  test <- data.frame(position = filtered_lollipop$pos, frequency = filtered_lollipop$AF, consequence = filtered_lollipop$HGVS)

  # Classify the consequences
  test <- test %>%
    mutate(variant_class = case_when(
      grepl(">", consequence) ~ "SNV",
      grepl("del", consequence) ~ "deletion",
      grepl("ins", consequence) ~ "insertion",
      grepl("dup", consequence) ~ "duplication",
      TRUE ~ "other"  # In case there are other types not specified
    ))
  
  # Define a color palette for the different variant classes
  variant_colors <- c("SNV" = "blue", 
                      "deletion" = "red", 
                      "insertion" = "green", 
                      "duplication" = "orange", 
                      "other" = "purple")
  
  variant_plot <- ggplot(test, aes(x=position, y=frequency, color = variant_class)) +
    geom_segment(aes(x=position, xend=position, y=0, yend=frequency), color="grey") +
    geom_point(size=2) +
    scale_color_manual(values = variant_colors) +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "top"
    ) +
    ggtitle(paste("Variant Frequencies of ", gene_name)) +
    ylab("Allele Frequency") +
    xlim(gene_start, gene_end)
  
  variant_plot_carpet <- variant_plot + 
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart1, xmax = orfend1, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart2, xmax = orfend2, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart3, xmax = orfend3, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart4, xmax = orfend4, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart5, xmax = orfend5, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart6, xmax = orfend6, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid")


  test_position <- data.frame(chromosome = lollipop$chrom, position = lollipop$pos, p = lollipop$AF, rsId = lollipop$filter)
  loc <- locus(data = test_position, gene = gene_name, fix_window = gene_end - gene_start,
             ens_db = ensDb_v106)
  grid.arrange(variant_plot_carpet, genetracks(loc, filter_gene_name = gene_name), nrow = 2, heights = c(1, 1))
}

```

The zoom_in_snp function creates a lollipop plot of the SNPs in the ORF of interest with a carpet plot of the corresponding DNA sequence
Input: 
- Name of the specific ORF of a lncRNA from the Ribo-seq_ORFs.csv file 

Output: 
- Lollipop plot of the SNPs in the ORF of interest with the allele frequency as the Y-axis and genomic position as the X-axis. The SNPs are distinguished by different colors based on their classification as SNV, insertion, deletion, duplication, delin, or other
- The exons of the ORF of interest is indicated by tall black boxes. The introns are left out of the graph
- A carpet plot of the corresponding DNA sequence of the exons is directly under the X-axis. Different nucleotides are represented by different colors

```{r}

zoom_in_snp <- function(orf_name){
  gene_name <- orf %>%
    filter(name2 == orf_name) %>%
    pull(geneName2)
  
  ensembl_name <-  orf %>%
    filter(name2 == orf_name) %>%
    pull(geneName)

  
  filtered_lollipop <- lollipop %>%
    filter(gene_id %in% ensembl_name)
  test <- data.frame(position = filtered_lollipop$pos, frequency = filtered_lollipop$AF, consequence = filtered_lollipop$HGVS)

  # Classify the consequences
  test <- test %>%
    mutate(variant_class = case_when(
      grepl(">", consequence) ~ "SNV",
      grepl("del", consequence) ~ "deletion",
      grepl("ins", consequence) ~ "insertion",
      grepl("dup", consequence) ~ "duplication",
      TRUE ~ "other"  # In case there are other types not specified
    ))
  
  # Define a color palette for the different variant classes
  variant_colors <- c("SNV" = "blue", 
                      "deletion" = "red", 
                      "insertion" = "green", 
                      "duplication" = "orange", 
                      "other" = "purple")
  
  linc_orf <- orf[orf$geneName2 == gene_name, ]

  rect <- data.frame(chromstart = linc_orf$chromStart, count = linc_orf$blockCount, size = linc_orf$blockSizes, start = linc_orf$blockStart)
  
  # Convert size and start columns to numeric vectors
  orf_df <- rect %>%
    mutate(size = map(strsplit(as.character(size), ","), function(x) as.numeric(x)),
           start = map(strsplit(as.character(start), ","), function(x) as.numeric(x)))
  
  # Function to create multiple columns based on count
  create_orf_columns <- function(chromstart, count, start, size) {
    # Initialize empty list to store results
    results <- list()
    
    # Loop over each count
    for (i in seq_len(count)) {
      # Calculate orfstart and orfend for each count
      orfstart <- chromstart + start[i]
      orfend <- chromstart + start[i] + size[i]
      
      # Store results in a list
      results[[paste0("orfstart", i)]] <- orfstart
      results[[paste0("orfend", i)]] <- orfend
    }
    
    # Return the results as a data frame
    return(as.data.frame(results))
  }
  
  # Apply the function row-wise using pmap from purrr
  orf_df <- bind_cols(orf_df, pmap_df(orf_df, create_orf_columns))
  orf_df$name <- linc_orf$name
  orf_df <- orf_df %>%
    mutate(ymax = ifelse(name == orf_name, Inf, 0))
  orf_df$chromEnd <- linc_orf$chromEnd
  
  iterations <- 8 - orf_df$count
    
  for (i in (orf_df$count + 1):iterations) {
    orf_df[[paste0("orfstart", i)]] <- 0
    orf_df[[paste0("orfend", i)]] <- 0
  }
  
  filtered_df <- orf_df %>% filter(name == orf_name)
  xmin <- min(filtered_df$chromstart)
  xmax <- max(filtered_df$chromEnd)
  
  variant_sizes <- c("SNV" = 2, 
                      "deletion" = 4, 
                      "insertion" = 4, 
                      "duplication" = 2, 
                      "other" = 2)
  

  filtered_row <- df[df$seq_name == orf_name, ]

  
  dna_df <- data.frame(Position = 1:nchar(filtered_row$sequence),
                           Nucleotide = strsplit(filtered_row$sequence, "")[[1]])
  
  sequences <- list()
  
  for (i in 1:filtered_df$count) {
    start_col <- paste0("orfstart", i)
    end_col <- paste0("orfend", i)
    sequences[[i]] <- seq(from = filtered_df[[start_col]] + 1, to = filtered_df[[end_col]], by = 1)
  }
  
  # Combine both sequences
  position <- unlist(sequences)

    # Create a data frame with the position column

  dna_df$Position <- dna_df$Position + xmin
  dna_df <- dna_df[dna_df$Position %in% position, ]
  
  nucleotides <- unique(dna_df$Nucleotide)
  colors <- rainbow(length(nucleotides))
  names(colors) <- nucleotides
  
  zoom_plot <- ggplot(test, aes(x=position, y=frequency, color = variant_class)) +
    geom_segment(aes(x=position, xend=position, y=0, yend=frequency), color="grey") +
    geom_point(aes(size=variant_class)) +
    scale_color_manual(values = variant_colors) +
    scale_size_manual(values = variant_sizes) +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "top"
    ) +
    ggtitle(paste("Variant Frequencies of ", orf_name)) +
    ylab("Allele Frequency") +
    xlim(xmin, xmax + 1) +
    geom_rect(data = dna_df, aes(xmin = Position - 0.5, xmax = Position + 0.5, ymin = -Inf, ymax = 0, fill = Nucleotide), color = "black", inherit.aes = FALSE) +  # rectangles below x-axis
    scale_fill_manual(values = colors) # fill colors for rectangles

  zoom_plot <- zoom_plot + 
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart1, xmax = orfend1, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart2, xmax = orfend2, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart3, xmax = orfend3, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart4, xmax = orfend4, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart5, xmax = orfend5, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart6, xmax = orfend6, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid")
  
    
  breaks <- list()
  num_orfs <- filtered_df$count
  
  # Loop through each ORF
  for (i in 1:(num_orfs - 1)) {
    breaks[[length(breaks) + 1]] <- filtered_df[[paste0("orfend", i)]] + 1
    breaks[[length(breaks) + 1]] <- filtered_df[[paste0("orfstart", i + 1)]] - 1
  }

  zoom_plot <- zoom_plot + scale_x_break(unlist(breaks))
  return(zoom_plot)
  
}

```

The protein_consequence function creates a lollipop plot of the protein consequence of the SNPs in the ORF of interest with a carpet plot of the corresponding DNA sequence. It also indicates the strand direction of the ORF
Input: 
- Name of the specific ORF of a lncRNA from the Ribo-seq_ORFs.csv file
- File with saved SNP consequence in variable called test

Output: 
- Lollipop plot of the SNPs in the ORF of interest with its severity as the Y-axis and genomic position as the X-axis. The SNPs are distinguished by different colors based on their protein consequence as missense_variant, stop_lost, synonymous_variant, frameshift_variant, and nonsense_variant
- The exons of the ORF of interest is indicated by tall black boxes. The introns are left out of the graph
- A carpet plot of the corresponding DNA sequence of the exons is directly under the X-axis. Different nucleotides are represented by different colors
- The strand direction is indicated using an arrow between the MODIFIER and LOW locations on the Y-axis

```{r}

protein_consequence <- function(orf_name){
  gene_name <- orf %>%
    filter(name2 == orf_name) %>%
    pull(geneName2)

  filtered_test <- test %>%
    filter(V4 %in% orf_name)
  filtered_test <- filtered_test %>%
    mutate(V2 = if_else(str_detect(V2, "-"),
                             str_extract(V2, "(?<=:)[0-9]+(?=-)"),
                             str_extract(V2, "(?<=:)[0-9]+")))
  filtered_test$V2 <- as.numeric(filtered_test$V2)
  filtered_test <- filtered_test %>%
    mutate(impact = str_extract(V14, "(?<=IMPACT=)[^;]+"))
  filtered_test <- filtered_test %>%
    mutate(strand = str_extract(V14, "(?<=STRAND=)[^;]+"))
  filtered_test <- data.frame(position = filtered_test$V2, orf_name = filtered_test$V4, Consequence = filtered_test$V7, impact = filtered_test$impact, strand = filtered_test$strand)
  
  impact_mapping <- c(
    "MODIFIER" = 0,
    "Strand Direction" = 0.5,
    "LOW" = 1,
    "MODERATE" = 2,
    "HIGH" = 3
  )
  filtered_test$impact_numeric <- impact_mapping[filtered_test$impact]


  
  # Define a color palette for the different variant classes
  variant_colors <- c("missense_variant" = "red", 
                      "stop_lost" = "orange", 
                      "synonymous_variant" = "green",
                      "frameshift_variant" = "blue",
                      "nonsense_variant" = "purple"
                      )
  
  linc_orf <- orf[orf$geneName2 == gene_name, ]

  rect <- data.frame(chromstart = linc_orf$chromStart, count = linc_orf$blockCount, size = linc_orf$blockSizes, start = linc_orf$blockStart)
  
  # Convert size and start columns to numeric vectors
  orf_df <- rect %>%
    mutate(size = map(strsplit(as.character(size), ","), function(x) as.numeric(x)),
           start = map(strsplit(as.character(start), ","), function(x) as.numeric(x)))
  
  # Function to create multiple columns based on count
  create_orf_columns <- function(chromstart, count, start, size) {
    # Initialize empty list to store results
    results <- list()
    
    # Loop over each count
    for (i in seq_len(count)) {
      # Calculate orfstart and orfend for each count
      orfstart <- chromstart + start[i]
      orfend <- chromstart + start[i] + size[i]
      
      # Store results in a list
      results[[paste0("orfstart", i)]] <- orfstart
      results[[paste0("orfend", i)]] <- orfend
    }
    
    # Return the results as a data frame
    return(as.data.frame(results))
  }
  
  # Apply the function row-wise using pmap from purrr
  orf_df <- bind_cols(orf_df, pmap_df(orf_df, create_orf_columns))
  orf_df$name <- linc_orf$name
  orf_df <- orf_df %>%
    mutate(ymax = ifelse(name == orf_name, Inf, 0))
  orf_df$chromEnd <- linc_orf$chromEnd
  
  iterations <- 8 - orf_df$count
    
  for (i in (orf_df$count + 1):iterations) {
    orf_df[[paste0("orfstart", i)]] <- 0
    orf_df[[paste0("orfend", i)]] <- 0
  }
  
  filtered_df <- orf_df %>% filter(name == orf_name)
  xmin <- min(filtered_df$chromstart)
  xmax <- max(filtered_df$chromEnd)
  
  variant_sizes <- c("missense_variant" = 4, 
                      "stop_lost" = 4, 
                      "synonymous_variant" = 2,
                      "frameshift_variant" = 4
                      )
  

  filtered_row <- df[df$seq_name == orf_name, ]

  
  dna_df <- data.frame(Position = 1:nchar(filtered_row$sequence),
                           Nucleotide = strsplit(filtered_row$sequence, "")[[1]])
  
  sequences <- list()
  
  for (i in 1:filtered_df$count) {
    start_col <- paste0("orfstart", i)
    end_col <- paste0("orfend", i)
    sequences[[i]] <- seq(from = filtered_df[[start_col]] + 1, to = filtered_df[[end_col]], by = 1)
  }
  
  # Combine both sequences
  position <- unlist(sequences)

    # Create a data frame with the position column

  dna_df$Position <- dna_df$Position + xmin
  dna_df <- dna_df[dna_df$Position %in% position, ]
  
  nucleotides <- unique(dna_df$Nucleotide)
  colors <- rainbow(length(nucleotides))
  names(colors) <- nucleotides
  
  zoom_plot <- ggplot(filtered_test, aes(x=position, y=impact_numeric, color = Consequence)) +
    geom_segment(aes(x=position, xend=position, y=0, yend=impact_numeric), color="grey") +
    geom_point(aes(size=Consequence)) +
    scale_color_manual(values = variant_colors) +
    scale_y_continuous(breaks = c(0, 0.5, 1, 2, 3), labels = c("MODIFIER", "Strand", "LOW", "MODERATE", "HIGH")) +
    scale_size_manual(values = variant_sizes) +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "top"
    ) +
    ggtitle(paste("Variant Consequence of ", orf_name)) +
    ylab("Impact Level") +
    xlim(xmin, xmax + 1) +
    geom_rect(data = dna_df, aes(xmin = Position - 0.5, xmax = Position + 0.5, ymin = -Inf, ymax = 0, fill = Nucleotide), color = "black", inherit.aes = FALSE) +  # rectangles below x-axis
    scale_fill_manual(values = colors) # fill colors for rectangles +


  zoom_plot <- zoom_plot + 
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart1, xmax = orfend1, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart2, xmax = orfend2, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart3, xmax = orfend3, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart4, xmax = orfend4, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart5, xmax = orfend5, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid") +
    geom_rect(data = orf_df,
              aes(x = NULL, y = NULL, xmin = orfstart6, xmax = orfend6, ymin = -Inf, ymax = ymax), 
              fill = NA, color = "black", size = 0.5, linetype = "solid")
  
    
  breaks <- list()
  num_orfs <- filtered_df$count
  
  if (num_orfs > 1) {
    # Loop through each ORF
    for (i in 1:(num_orfs - 1)) {
      breaks[[length(breaks) + 1]] <- filtered_df[[paste0("orfend", i)]] + 1
      breaks[[length(breaks) + 1]] <- filtered_df[[paste0("orfstart", i + 1)]] - 1
    }
    zoom_plot <- zoom_plot + scale_x_break(unlist(breaks))
  }
  
  if (any(filtered_test$strand == 1)) {
    zoom_plot <- zoom_plot + 
      geom_segment(
        aes(
          x = (xmin + xmax) / 2 - 10,
          xend = (xmin + xmax) / 2 + 10,
          y = 0.5,
          yend = 0.5
        ),
        arrow = arrow(type = "closed", length = unit(0.2, "inches")),
        size = 1,
        color = "black"
      )
  }
  
  # Add arrow pointing left if strand is not 1
  if (any(filtered_test$strand != 1)) {
    zoom_plot <- zoom_plot + 
      geom_segment(
        aes(
          x = (xmin + xmax) / 2 + 10,
          xend = (xmin + xmax) / 2 - 10,
          y = 0.5,
          yend = 0.5
        ),
        arrow = arrow(type = "closed", length = unit(0.2, "inches")),
        size = 1,
        color = "black"
      )
  }
    
  return(zoom_plot)
}

```

---------Main---------------
0. Install and call all the packages at the start of the script
1. Run through the code to load necessary lncRNA, SNP, and sequence data
2. Specify your ORF of interest and call the functions zoom_out_tracks, track_with_snp, and zoom_in_snp to visualize the SNP data
3. If you have a file with protein consequence, read the file into a data frame and save it to the variable "test". Specify your ORF of interest and call the function proetin_consequence to visualize the protein consequence of the SNPs

```{r}
#-----------------------------------Load lncRNA data and name the columns------------------------------------------------
orf <- read.csv("Ribo-seq_ORFs.csv", header = F, sep = "\t")
names <- c("chrom", "chromStart", "chromEnd", "name", "score", "strand", "thickStart", "thickEnd", "itemRgb", "blockCount", "blockSizes", "blockStart", "name2", "cdsStartStat", "cdsEndStat", "exonFrames", "type", "geneName", "geneName2", "geneType", "transcript_biotype", "proteinSequence", "all_transcript_ids", "all_gene_ids", "replicated", "ref_studies")
colnames(orf) <- names


#-----------------------------------Load SNP data and add column called HGVS that classifies the SNPs--------------------
lollipop <- read.csv("variants_gencode_orfs.csv")

create_hgvs <- function(seqname, start, ref, alt) {
  if (nchar(ref) == 1 && nchar(alt) == 1) {
    # SNV
    return(paste0(seqname, ":c.", start, ref, ">", alt))
  } else if (nchar(ref) == 1 && nchar(alt) > 1) {
    # Insertion
    return(paste0(seqname, ":c.", start, "_", as.numeric(start) + 1, "ins", alt))
  } else if (nchar(ref) > 1 && nchar(alt) == 1) {
    # Deletion
    if (nchar(ref) == 1) {
      return(paste0(seqname, ":c.", start, "del"))
    } else {
      return(paste0(seqname, ":c.", start, "_", as.numeric(start) + nchar(ref) - 1, "del"))
    }
  } else if (nchar(ref) > 1 && nchar(alt) > 1 && ref != alt) {
    # Indel
    return(paste0(seqname, ":c.", start, "_", as.numeric(start) + nchar(ref) - 1, "delins", alt))
  } else if (ref == alt) {
    # Duplication
    if (nchar(ref) == 1) {
      return(paste0(seqname, ":c.", start, "dup"))
    } else {
      return(paste0(seqname, ":c.", start, "_", as.numeric(start) + nchar(ref) - 1, "dup"))
    }
  } else {
    return(NA)
  }
}

lollipop$HGVS <- apply(lollipop, 1, function(row) {
  create_hgvs(row["chrom"], row["pos"], row["ref"], row["alt"])
})


#-----------------------------------Load sequence data of each lncRNA orf and remove extra information from the first column-----------------------
fastaFile <- readDNAStringSet("Ribo-seq_ORFs.fa")
seq_name = names(fastaFile)
sequence = paste(fastaFile)
df <- data.frame(seq_name, sequence)
df$seq_name <- sub(":.*", "", df$seq_name)


#-----------------------------------Call the functions to graph the plots---------------------------------------
zoom_out_tracks("c1riboseqorf1")
track_with_snp("c1riboseqorf1")
zoom_in_snp("c1riboseqorf1")
  
#-----------------------------------Visualize Protein Consequence
test <- read.csv("vep_results_ncORFs_test.csv", header = F)
protein_consequence("cYnorep1")

```

gene_symbol <- "LINC01128"  # Replace with your gene of interest

# Use the Ensembl BioMart to fetch the gene sequence
mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# Get all exons of all transcripts for the gene
gene_sequence <- getSequence(id = gene_symbol,
                             type = "hgnc_symbol",
                             seqType = "gene_exon",
                             mart = mart)


# Combine all sequences (in case the gene has multiple transcripts)
dna_sequence <- paste(gene_sequence$gene_exon, collapse = "")

#Get all transcripts
transcripts <- getBM(
  attributes = c('ensembl_transcript_id', 'transcript_biotype', 'transcript_start', 'transcript_end', 'transcript_length'),
  filters = 'hgnc_symbol',
  values = gene_symbol,
  mart = mart
)

# Filter to select only the principal transcript (you may need to adjust this based on your criteria)
# Here, we are selecting the longest transcript as a proxy for the principal transcript
principal_transcript <- transcripts[which.max(transcripts$transcript_length), 'ensembl_transcript_id']

# Get the exon sequences for the principal transcript
exon_sequences <- getSequence(
  id = principal_transcript,
  type = "ensembl_transcript_id",
  seqType = "gene_exon",
  mart = mart
)

--Victor Wang 2024/7/16 :D

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

